## 1. Walleye pollock


Status and catch specifications (t) of pollock and projections for `r year+1` and `r year+2`. 
Biomass for each year corresponds to the projection given in the SAFE report issued in the preceding year (age 3+ for W/C/WYAK and survey biomass for SEO). 
The OFL and ABC for `r year+1` and `r year+2` are those recommended by the Plan Team. 
Catch data were through `r catch_data_date`. 

The GOA-wide and W/C/WYAK ABCs listed in this table are before reductions for the Prince William Sound GHL. 
However, the federal TACs from earlier years reflect reductions from the ABC due to State waters GHL. 
State waters GHL was computed as 2.5% of the total W/C/WYAK ABC.

```{r}
spp = "pollock"

specs %>% 
  dplyr::filter(stock %in% spp) %>% 
  dplyr::mutate(area = case_when(area == 'Total' ~ "Goa-wide",
                                 area %in% c('EYAK/SEO') ~ "SEO",
                                 TRUE ~ "W/C/WYAK")) %>% 
  group_by(area, year, measure) %>% 
  summarise(value = round(mean(value, na.rm = T))) %>% 
  tidyr::pivot_wider(names_from = measure, values_from = value) -> specs2
  
 bio %>% 
    dplyr::filter(stock %in% spp) %>% 
    dplyr::rename(!!planteam::nameit(.) := biomass) %>% 
    dplyr::select(-type, -stock) %>% 
    left_join(specs2, .) %>% 
    dplyr::select(Area = area, Year = year, dplyr::contains("biomass"), 
                  OFL=ofl, ABC=abc, TAC=tac, Catch=catch) %>% 
    dplyr::arrange(dplyr::desc(Area)) %>%
    flextable::flextable() %>% 
    flextable::merge_v(j=1) %>% 
    flextable::colformat_num(j=1, big.mark = "") %>% 
    flextable::hline(i=4) %>% 
    flextable::align(j=1, part="all") %>% 
    flextable::valign(valign="bottom", part="header") %>% 
    flextable::fix_border_issues() 

```



*Changes from the previous assessment*
This year’s pollock assessment features the following new data: 
1) 2020 total catch and catch-at-age from the fishery, 
2) 2021 biomass and age composition from the Shelikof Strait acoustic survey, 
3) 2021 NMFS Bottom Trawl survey biomass and length composition, 
4) 2021 Summer GOA-wide acoustic survey biomass and length composition, 
and 5) 2021 biomass and 2020 age composition from the ADF&G crab/groundfish trawl survey.
The age-structured assessment model used for GOA W/C/WYAK pollock assessment was identical to the 2019 and 2020 assessments (Model 19.1).

*Spawning biomass and stock trends*
The spawning stock is projected to continue to decline slightly in 2022 and 2023 as the 2012 year class is
further reduced in abundance, however with new fish recruited into the fishery, spawning biomass is
projected to start increasing in 2025. The presence of several incoming year classes should result in a
stabilization in biomass. However, the 2021 Shelikof Strait survey showed an unexpected reduction in the
estimated abundance of the 2018 year class (aged 3) relative to their abundance in the 2019 survey. The
2017 year-class (aged 4) is still present in high numbers, and a strong new 2020 class was detected in all
surveys. Overall, the Shelikof Strait survey data in 2021 showed a similar biomass to 2020. Overall,
survey indices seem to be providing similar trends with closer agreement with the ADF&G survey, as
well as the 2021 NMFS bottom trawl survey results. An exception to this was the acoustic summer
survey, which was 25% percent lower than the 2019 estimate.

*Tier determination/Plan Team discussion and resulting ABCs and OFLs*
The model projection of female spawning biomass in 2022 is 186,481, which is above B40% (172,000),
which places the W/C/WYAK Gulf of Alaska pollock stock in Tier 3a. 
The model estimated 2022 age-3+ biomass is 1,097,340 t (for the W/C/WYAK areas) which was similar to the 2020 estimate (1,007,850 t). 
The author scored the current risk conditions as Level 1 for all four risk categories, and thus did not
recommend a reduction from maximum permissible ABC.
The Team supported the authors’ recommendation using the assessment-derived maximum permissible
ABC for 2022. The resulting 2022 ABC for pollock in the Gulf of Alaska west of 140° W longitude
(W/C/WYAK) is 133,081 t which is a 26 % increase from the 2021 ABC. 
The OFL is 154,983 t for 2022.
The 2021 Prince William Sound (PWS) GHL is 3,327t (2.5% of the W/C/WYAK ABC).
Pollock in southeast Alaska (East Yakutat and Southeastern areas) are in Tier 5. 
The recommended ABC is 11,363 t for 2022 and 2023, which is an increase of 12% from the 2020 ABC. 
These recommendations are based on natural mortality (0.3) and the random effects model fit to the 1990-2021 bottom trawl survey biomass estimates in Southeast Alaska.

*Status determination*
The Gulf of Alaska pollock stock is not being subjected to overfishing and is neither overfished nor
approaching an overfished condition.

*Area apportionment*
The assessment was updated to include the most recent data available for area apportionments within each
season (Appendix 1D.3 of the GOA pollock chapter). 
For winter seasons, model estimates of biomass for winter acoustic surveys conducted were used as a basis for apportionment. 
pportionments for the B1 and B2 seasons were based on a 3-year weighted average of the sum of the AFSC bottom trawl survey and the
gulf-wide acoustic summer survey (unchanged from the previous assessment). 
Area apportionments, including the 2.5% of the ABC for the State of Alaska managed pollock fishery in Prince William Sound, are as follows:

```{r}
app %>% 
  dplyr::filter(stock %in% spp) %>% 
  tidyr::pivot_wider(names_from = area, values_from = value) %>% 
  dplyr::select(Year=year, 4:8, "PWS GHL" = `State GHL`, Total) %>% 
  flextable::flextable() %>% 
  flextable::colformat_num(j=1, big.mark = "") %>% 
  flextable::align(j=1, part="all") %>% 
  flextable::valign(valign="bottom", part="header")
```

