## 13. Rougheye and blackspotted rockfish
Status and catch specifications (t) of rougheye and blackspotted rockfish and projections for `r year+1` and `r year+2`. 
Biomass for each year corresponds to the projections given in the SAFE report issued in the preceding year. 
The OFL and ABC for `r year+1` and `r year+2` are those recommended by the Plan Team. 
Catch data are current as of `r catch_data_date`. 

```{r}
spp = c("rougheye and blackspotted rockfish", "rebs")

specs %>% 
  dplyr::filter(stock %in% spp, area=="Total") %>% 
  tidyr::pivot_wider(names_from = measure, values_from = value) -> specs2
  
sum_tbl(spp, bio, specs2)
```

*Changes from the previous assessment*
This year was a full assessment and the authors used updated catch data, trawl and longline survey biomass, and fishery and longline ages. 
There have been no model changes for this assessment since 2015.

*Spawning biomass and stock trends*
Estimated female spawning biomass for 2022 is 8,648 t. 
This is above the $B_{40\%}$ value of 5,911 t.

*Tier determination/Plan Team discussion and resulting ABCs and OFLs*
The rougheye/blackspotted complex qualifies as a Tier 3 stock. 
For 2022 and 2023, the Plan Team accepted the authors’ recommended maximum permissible ABCs and the OFLs as provided in the table above.

*Status determination*
The stock is not being subject to overfishing, is not currently overfished, nor is it approaching a condition of being overfished.

*Area apportionment*
The recommended apportionments for 2022 and 2023 are calculated using the two-survey random effects model, which was approved for use in this assessment in 2019. 
This method equally weights the longline and trawl survey indices.
Rougheye and blackspotted rockfish area apportionment for ABCs are shown below:

```{r}
app %>% 
  dplyr::filter(stock %in% spp) %>% 
  tidyr::pivot_wider(names_from = area, values_from = value) %>% 
  dplyr::mutate(Total = sum(W, C, E)) %>% 
  dplyr::select(Year=year, "Western" = W, "Central" = C, "Eastern" = E, Total ) %>% 
  flextable::flextable() %>% 
  flextable::colformat_num(j=1, big.mark = "") %>% 
  flextable::align(j=1, part="all") %>% 
  flextable::valign(valign="bottom", part="header")
```
